### 2015-02-27 project1/checkpoint/main.tex
Adding a summary section a few citations

\documentclass[a4paper,12pt]{article}

\usepackage{url, hyperref, graphicx}

\begin{document}

\title{Project 1 Checkpoint}
\author{Chip Bell}
\date{February 6th, 2015}
\maketitle

\section{Problem Description}
Traffic simulation is becoming increasing important due to the expansion and growth of cities. Long-standing
infrastructure is forced to adapt in order to accomodate greater amounts of traffic, despite the fact that in many
cases the roads cannot be modified. This can be a result of legislation, existing buildings, and private property among
other things. So as a result, cities must instead optimize the variables they can control. In this case, we consider
signal light timing for the 12th St. and Peachtree Street intersection. In particular,
we explore the effects that light timing has on traffic flow, given a source input distribution of cars.

\section{The Conceptual Model}
In order to create a representative software simulation of the aforementioned system, we first begin by designing a
conceptual version of this model from which we can plan our implementation. We'll borrow heavily from the results and data
in \cite{cts12} when formulating our model.
Our architecture will be event driven, prioritizing events by timestamp. Furthermore,
we'll utilize the publisher-subscriber pattern (see \cite{pubsub}) to promote better decoupling.
Given this architecture, our first step will be to identify the entities present in our model, and then consider their
interactions (or the inherent ``activies'') within the simulation. This will allow us then to design the events that
are important and relevant to the simulation.

\subsection{Entities}
The first entity, unsurprisingly, is a single car. Our representation of a car is clearly less detailed than the real
thing, but rather focuses on variables that influence it's travel time from its starting location to destination, along
with the variables that influence neighboring cars. For our simulation, we'll track the position and destination of the
cars. Probably the most primary value associated with a car is its arrival time, since that dictates when and where it
will fall in the queue of cars at the intersection.

Another entity to consider is the intersection itself. It's changing in time constantly, and those changes affect
every entity within the simulation. The most important parameters to consider here are the light timings. This
intersection's timings are time-based, rather than sensor-based, which can be seen in the data set provided with the
project description and is also available online \cite{ngsim}. We assume that the current light timing provided here
are optimal, but because we'll have a computer simulation in the end we have free reign in experimenting with these
values and observing the outcome. Perhaps the timings are \emph{not} optimal. Perhaps, our model grossly oversimplifies
some crucial interaction within the system. This is perhaps the most exciting part of the project.

\subsection{Basic Interactions}
From the intersection entity, we can glean an important interaction that occurs: The changing of light signal. When a
signal changes, each car in-queue will be forced to change it's state, whether it's waiting or currently in motion.
From an event-driven perspective, this means that any listener would need to be able to know \emph{how} the light
changed in order to act accordingly.

An intersection serves as holding point for cars traveling along the road so if a light is red, cars are forced to
queue up at the intersection until green. This lends itself to a queue-style data structure that encorporates both
position and acceleration. Car acceleration is a complex topic, but models have been constructed \cite{bonneson}
\cite{herman_et_al}. These models are generally continuous, and can generally be represented as a differential equation
due to the relationships between acceleration and velocity, as can be seen in \cite{briggs} \cite{deceleration} as well.
Instead, we will simply consider the various events that trigger changes in behavior. For instance, a real driver is
constantly observing their environment (hopefully) and making minute adjustments the entire time. However, we'll take
a greatly simplified approach and simply precalculate an estimate of how long a car takes to traverse the section of
road, along with a model for accelerating from a stop.

\subsubsection{Acceleration Model}
Bonneson \cite{bonneson} noted a property of car platoon acceleration: Acceleration decreases linearly with velocity.
The formulation he provides is a differential equation:
\[
\frac{dV}{dt} = A_{m} \left( 1 - \frac{V}{V_{m}} \right)
\]
Where $A_m$ and $V_m$ where the empirically calculated maximum acceleration and velocity of cars. The acceleration was
determined to be roughly 6.63 $\frac{ft}{s^2}$ and the maximum velocity simply the speed limit in the area (30 mph).
This equation is indeed solvable by hand and gives us a closed form for the position of car (assuming an initial velocity of 0):
\[
    p(t) = V_{m}t - \frac{V_{m}^2}{A_{m}} \left( 1 - e^{-\frac{A_{m}t}{V_{m}} } \right)
\]

Assuming all cars have length $L$, the $n$th car in the queue will have to travel $L(n-1)$ feet in order to cross the
intersection. Given our closed form above, in order to know how long it will be before a car passes we'll need to solve
\[
p(t) = L(n-1)
\]
Unfortunately, this must be done numerically. The function is fairly well behaved, and will in general have two roots
(only one positive). Using a positive initial guess (around 2 or so, judging by visual inspection), Newton's method
will converge within only a few steps. \ref{fig:acceleration} demonstrates some properties of this function
$0 = p(t) - L(n-1)$, one being
that as we move further back in the queue, the timing between cars approaches a constant. This is a result of cars
have more space to accerate to full speed.

\begin{figure}
\begin{center}  
\includegraphics[width=4in]{../acceleration.pdf}  
\caption{\small \sl $p(t) - L(n-1)$ for $n \in \{0, 1, 2, 3, 4\}$ \label{fig:acceleration}}  
\end{center}  
\end{figure}

Given this formulation, we can use this method to calculate when a given car will pass through the intersection,
assuming it was at a complete stop. We'll use this information when enqueueing the "exit" event for a car.

\subsection{Events}
Based on our observations from above, we can now construct a set of events that are pertinent to this application.

First off, signal changes consititute the major event in the system. When a signal changes, the event must contain
\emph{when} the signal changed, and \emph{what} the signal changed to. The light itself will listen for this event
and will enqueue the next signal change based on the signal change time. These signal changes could be enqueued
completely upfront, but allowing this to be done on the fly allows the user to tweak these values during the simulation.

The car model will listen for this event as well. When a currently traveling car sees a red light, it will calculate
\emph{where} it needs to stop, and \emph{when} it will come to a stop. This will be done using a deceleration model
similar to \cite{deceleration}. If the light turns to green, the car instead needs to start accelerating (using one of
the models mentioned above). The model will adapt to light changes during movement by recalculating its estimated
exit time of the system.

\subsection{Simulation Input}
The two main sources of input into this simulation are the cars entering out section of Peachtree St, and the light
timing of the system. The combination of these two variables are what influence the smoothness
of traffic flow to the actual drivers in the system.

We first concern ourselves with finding an input distribution for cars entering the system. In our model, we'll only
consider basic aspects of a car: when it enters the system, and it's intended destination (north or south). We will assume that all cars have the same speed entering the system, and that they
have the same length. These values can be calculated from the NGSIM data set \cite{ngsim} (which yields an average car length of 16.407).

\begin{figure}
\begin{center}  
\includegraphics[width=4in]{../northbound.png}  
\caption{\small \sl Northbound Arrival Intervals by Order of Occurrence\label{fig:northbound_arrival}}  
\end{center}  
\end{figure}

\begin{figure}  
\begin{center}
\includegraphics[width=4in]{../southbound.png}
\caption{\small \sl Southbound Arrival Intervals by Order of Occurrence\label{fig:southbound_arrival}}  
\end{center}  
\end{figure}

The distribution of car arrival is very important, but has been difficult to model. The arrival intervals between 
northbound and southbound cars are provided in \ref{fig:northbound_arrival} and \ref{fig:southbound_arrival}.
Model fitting was attempted for both $\Gamma$ and $\Gamma$-mixture models on the raw data, but resulted in poor
fits. Furthermore, pruning outliers was attempted as well, but still yielded bad matches. The outlier critera used for
this determination is the common interquartile range approach, where values are rejected when
they lie outside of $median + 1.5IQR$, where $IQR$ is the interquartile range $Q_3 - Q_1$.

Given the difficulties with modeling we'll take a simpler approach and assume the ``true distribution`` is modeled by what was observed,
and simply plan to use the observed frequency distribution as our input distribution.

The current light timing of the system is known, and provided in \cite{ngsim}, but we'll allow the user to set
those values dynamically via a UI.

\subsubsection{Data Analysis}
In the \texttt{data\_processing} folder, most of code used for distribution fitting and statistics can be found there.
The data analysis code was written in JavaScript, and runs completely in NodeJS \cite{nodejs}. NPM \cite{npm} was used
to manage dependencies and those dependencies can be seen in the \texttt{package.json} file. Of those packages, I have
authored the \texttt{chi-squared-test} module \cite{chiSquaredNpm}, and the \texttt{gamma-distribution}
\cite{gammaDistributionNpm}. Gulp \cite{gulp} was used a build tool for automating tasks (including building this paper).
To run some of the data analysis code I wrote, you'll need to install and run gulp. Instructions for doing so are
provided in the final section of this paper.

Alongside the NGSIM data set \cite{ngsim}, we are also dependent on the browser for much of our runtime configuration.
In particular, we rely on the random number generation capabilities of the browser for our entire source of randomness.
JavaScript provides a built-in implementation \cite{ecmascriptmath} \texttt{Math.random()}, which provides a uniformly
distributed random number between 0 and 1. The standard also specifies that the implementation is vendor-dependent.
This means that the "randomness" of \texttt{Math.random()} is dependent on the browser. Because of this, I've provided
a web page to calculate a simple $\chi^2$ statistic for a sample \cite{rngtester}. I've tested on Safari, Firefox, and
Chrome with generally solid results, despite the fact that their implementations are different
\cite{rngimplementation}. However, Hyland \cite{rngimplementation} does warn the reader about the caveats associated
with the native browser implementation of random numbers. However, these issues are not terribly important for a
simulation of this simplicity.

\subsection{Simulation Output}
The most important statistic we'll want to collect during the simulation is the thoughput of the simulation. We'll
want to observe how the light timing affects the average travel time for cars. It will also to see how variable this
is. For instance, do cars travelling on Peachtree St. have similar travel times, despite their arrival time, or does
their arrival drastically lengthen their time to travel the stretch of road we are modeling. Furthermore, we might also
ask if the light timing has that much of an effect at all, and if it does what is the ``critical point`` at which the
system no longer can serve incoming cars fast enough?

\subsection{Simplifications and Assumptions}
Probably the most obvious "flaw" in such a design are the simplifications made. Details can certainly be left out
purposely for many reasons such as time constraints, lack of supporting data to model a extra feature, or simply
unneccesary complexity.

In this case, certainly the most prominent would be the ``isolation'' of the model. We are only considering an
idealized random input distribution of cars from both North and South of the 12th St. intersection. However, this is
certainly not the case, since this input distribution is affected by countless other factors like the surrounding
lights, cars passing along 12th at that intersection, and even simply time of day. We assume in this simulation is
perpetually between 4:00PM to 4:15PM in Atlanta (a frightening thought indeed), simply due to the fact that we only
have data for that time range. All in all, we assume the conditions were fairly identical to that in which the 
aforementioned CTS study was made.

Another simplification to consider is lane changing. Lane changing models do exist \cite{lanechanging} \cite{lanechanging2},
but can be complex, requiring modeling of driver choices and gap acceptance for drivers. Peachtree is a two lane road,
so lane changes are possible. However, we will make the assumption that drivers by this point have already made any
discretionary lane changes and traffic is evened out between lanes. This allows us to consider only a single lane at a
time, since the second lane is identical to the first after these assumptions.

We also assume here that cars are entering at the same velocity, based off of the urban speed limit of 30 mph set by
the state of Georgia \cite{gacode}. Furthermore, we assume a single sized car, based off of the average from the NGSIM data.
Having a constant speed and length allows us to simplify the calculation of stop points for cars entering the
intersection. Although these simplifications help, dealing with acceleration is an even more complex matter. In this
case it is both a mixture of both discrete (the light signal) and continuous elements (the car's velocity). By choosing
an event-driven approach we are forced to simplify acceleration, since we cannot accurately represent a continuous
system in this manner.

Due to our inability to map a common distribution to the data gathered in the NGSIM study, we are using an empircal
distribution to model the input distribution of cars. Although this is an "easy" choice, it comes with the assumption
that both the measured data is representative of the true input distribution of cars, and that there is no measurement
error in the data collected.

\section{Simulation Software Architecture}
Because of the event-driven architecture, the system can be though of as a set of event emitters and event listeners.
In some cases, an object may both emit and listen for events.

The first event emitter to consider is the \texttt{CarEmitter}. The \texttt{CarEmitter} uses the measured input
distribution we have for cars and generates a set of cars randomly. It then emits these cars as individual
\texttt{car:arrived} event. There will generally be two \texttt{CarEmitter}s, one for the North and South direction.

The second emitter will be the \texttt{LightSignal}. It will emit \texttt{light:changed} events, passing the light color.
Also, instead of enqueueing these events upfront, it will also listen for \texttt{light:changed} so that its
changing duration can modified on the fly (using the UI).

The \texttt{IntersectionQueue} holds a list of cars lined up at the intersection. It also maintains the
current signal state, so that it can send cars directly through the intersection if the light is green. It responds
to pretty much every event mentioned above. When the light changes, it will need update it's current signal state.
This may mean sending some cars off of the queue and stop holding cars, or to no longer let cars through and make them
wait. Furthermore, when a car arrives it will handle that event by deciding whether the car can pass through or needs
to wait on the light. Also, this class emits events. Whenever a car finally crosses the intersection, it emits
a \texttt{car:exited} event.

Since we want to gather metrics regarding the throughput of cars, we'll also provide a listener class
\texttt{ThroughputStatistics} that will listen for cars to complete their journey through the system. This will
be reported in the UI in real time.

The simulation is written in CoffeeScript \cite{coffeescript}, which is a small language that compiles to JavaScript.
CoffeeScript removes some of the boilerplate associated with setting up prototype chains and other object-oriented
concepts in JavaScript.
Backbone \cite{backbone} along with Underscore \cite{underscore} and jQuery \cite{jquery} were used to provide some
application structure, like basic eventing, and the sorted
collections that are used to form the basis for the event queue and \texttt{IntersectionQueue}. Lastly, Browserify
\cite{browserify} was used to bundle the code into a single file. Minification was done with Uglify.js \cite{uglify}.
Graphing is performed via inline SVG \cite{svg}.

\section{Current Build and Running Instructions}
The simulation is currently deployed at \cite{aprilandchip} and provides a UI that allows the light change intervals to
be updated in real time. Moreover, the simulation speed can be adjusted in real-time as well. Worth noting in the UI is
a log of all events that are passing through the system. This is live filterable. Also, a visualization of the current
signal state, along with the current wait times for the first 5 cars in line is provided. Statistics are provided in
real-time, providing both the observed input distribution of car inter-arrival times, and the output distribution of
car wait times.

To build the application locally, you'll need to install a version of Node.js \cite{nodejs} at 0.10.0 or higher. This will be bundled
with npm. Since this project uses gulp for a task runner, you'll need to install it globally via
\texttt{npm install -g gulp}. Inside the \texttt{simulation} directory, you'll need to pull in packages with just
\texttt{npm install}. Once that installs, running \texttt{gulp} will run the build for you and generate a
\texttt{main.js} that's included in the \texttt{index.html} file. Opening this file in a browser should run the
simulation. To rerun the simulation, simply refresh the browser

\section{Results}
Since light timing is adjustable in real-time, the actual average wait times can vary (even disregarding randomness).
However, a few general trends tend to appear. First off, the distribution of wait times is bimodal. This can be
somewhat observed in the NGSIM data set. This appears to be a result of the dichotomy of cars arriving when the light
is green, versus when it is red. Cars that arrive during a green light in general do not have to wait long at all
before they can cross the intersection. However, cars that arrive during a red light will be forced to wait, often
behind other cars. This causes their wait time to be much higher since they actually had to queue up before they could
exit the system. Unsurprisingly, most cars that catch the red light end up waiting around the duration of red time.

A second observation is the system saturation. There is a point where the light timing is so terrible that the system
cannot push cars through as fast as they come. In queueing theory terms, the system is referred to as being
``oversaturated``. After tweaking some values, this can be triggered by setting the red duration to 60 seconds and
slowing adjusting the yellow and green down (my testing resulted in oversaturation around yellow and green with 1 and
25 second durations respectively).

\section{Conclusion}

\bibliographystyle{plain}
\bibliography{main}

\end{document}